---
title: "notes"
author: "John Hinkle"
date: "2025-01-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ADA Notes

### Load in {tidyverse} and data

```{r}
library(tidyverse)
f <- "/Users/jeh5663/Documents/data-wrangling.csv"
d <- read_csv(f, col_names = TRUE)
```

### Exploring the data

#### Class

```{r}
class(d)
```

#### Structure

What type of data is in each column?

```{r}
str(d)
```

#### Glimpse

Similar to str(), but a bit cleaner presentation

```{r}
glimpse(d)
```

#### Head

First 5 lines of data

```{r}
head(d, 5)
```

#### Tail

Last 5 lines of data

```{r}
tail(d, 5)
```

#### Group Size Data Exploration

Method 1 (Vector)

```{r}
gs <- d$MeanGroupSize
class(gs)
head(gs)
str(gs)
```

```{r}
head(gs + 1) # Adds 1 to each group size

#na.rm = TRUE removes NA/missing values 
mean(gs, na.rm = TRUE) # mean
sd(gs, na.rm = TRUE) # standard deviation
```

```{r}
sciname <- paste0(d$Genus, " ", d$Species)
sciname <- toupper(sciname) # change all letters to uppercase
sciname
```

```{r}
unique(d$Genus) # pull out all unique values
```

Method 2 (Table)

```{r}
gs <- d[, "MeanGroupSize"]
class(gs)
head(gs)
str(gs)
```

### In-Class Exercise

#### Load {tidyverse} and data

```{r}
library(tidyverse)
f <- "/Users/jeh5663/Documents/data-wrangling.csv"
d <- read_csv(f, col_names = TRUE)
```

#### Create female body mass and brain size objects

```{r}
bm <- d$Body_mass_female_mean
bs <- d$Brain_Size_Species_Mean
```

#### Un-transformed plot

```{r}
plot(bm, bs)
```

#### Log transformed plot

```{r}
plot(log(bm), log(bs))
```
